### 2019-11-15 comp_model_1019.Rmd
added tabular Q learning

---
title: "opponent modelling"
author: "Maarten Speekenbrink"
date: "24 July 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## Reading data

```{r}
dat <- read.csv("data20180719.csv")
#dat <- read.csv("test_data.csv")
```

## Preprosessing and transforming data

```{r}
# using some functions from the 'tidyverse' (also for me to get use to them ;-)
library(tidyr)
library(dplyr)
library(DEoptim)
# transform 'winner' variable in numeric score
dat$score <- recode(dat$winner, human = 1, tie = 0, ai = -1)
# create a new variable 'block' with round 1...25 = block 1 and round 26...50 as block 2
dat$block <- as.numeric(cut(dat$round,2))
# create a new data.frame in a 'wide format'
# widedata <- dat %>%
#   unite(game_block,game,block) %>% # this creates a new variable which combines game and block
#     group_by(human_id,condition,game_block) %>% # let the functions know you want to separate things by ID, condition, and game_block
#       summarize(mean_score = mean(score)) %>% # compute average score (i.e wins - losses)
#         spread(game_block,mean_score) # reformat in the 'wide' format
# # save the data file as a .csv to use in e.g. SPSS
# write.csv(widedata,row.names=FALSE,file="scores_wide.csv")

```

```{r estimate-models, cache=TRUE}
# read in the various files which 
rps_predict_opp <- read.csv("rps_predict_opp.csv")
fwg_predict_opp <- read.csv("fwg_predict_opp.csv")
numbers_predict_opp <- read.csv("numbers_predict_opp.csv")

# recode actions to make them equal to the codes in these files
dat$h_action <- recode(dat$human_action,"rock" = "R", "paper" = "P", "scissors" = "S", "fire" = "F", "water" = "W", "grass" = "G", "one" = "1", "two" = "2", "three" = "3", "four" = "4", "five" = "5")
dat$a_action <- recode(dat$ai_action,"rock" = "R", "paper" = "P", "scissors" = "S", "fire" = "F", "water" = "W", "grass" = "G", "one" = "1", "two" = "2", "three" = "3", "four" = "4", "five" = "5")

# logit tranformation
my_logit <- function(x) {
  ret <- log(x/(1-x))
  ret[x==0] <- -Inf
  ret[x==1] <- Inf 
  return(ret)
}

# inverse logit transformation
my_logistic <- function(x) {
  ret <- 1/(1+exp(-x))
  ret[x == -Inf] <- 0.0
  ret[x == Inf] <- 1.0
  return(ret)
}

```

## Modelling fixed type participant
```{r}
# FIXED STRATEGIES (level-k) 
# create a function which computes the -2 log Likelihood over the three games for a particular strategy
# modelLL <- function(par,data,strategy, return_value=c("-2logLik","likelihood_by_trial")) {
#   # par = parameter vector (just has one element here) 
#   # dat = data subset for one participant
#   # strategy = c("level1","level2","level3","WSLS")
#   # return = -2logLik (default, for use in optimization) or "likelihood_by_trial" (for later plotting, etc)
#   return_value <- match.arg(return_value)
#   
#   # count the number of supplied parameters. If 1, it is a _1 model, 
#   # if 3, it is a _3 model. For the _3 model we will use unbounded
#   # optimization (transforming parameters from a logit scale back to 0 to 1)
#   n_theta <- length(par)
#   if(n_theta == 3) par <-my_logistic(par)
#   theta <- par[1]
#   # set up a vector to store the likelihood
#   lik <- rep(0.0,nrow(data))
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     pred_file <- switch(as.character(t_game),rps=rps_predict,fwg = fwg_predict, numbers = numbers_predict)
#     if(n_theta == 3) {
#       theta <- switch(as.character(t_game),rps=par[1],fwg = par[2], numbers = par[3])
#     }
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       pred <- NA
#       lik[t] <- 1/nopts
#     } else {
#       # get prediction from file
#       pred <- as.character(filter(pred_file,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
#       if(as.character(data[t,"h_action"]) == pred) {
#         lik[t] <- theta + (1-theta)/nopts
#       } else {
#         lik[t] <- (1-theta)/nopts
#       }
#     }
#     # for debugging uncomment the line below
#     # cat(as.character(pred),as.character(data[t,"h_action"]),lik[t],"\n")
#   }
#   if(return_value == "-2logLik") {
#     ret <- -2*sum(log(lik))
#     if(ret == Inf) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik)
# }
# 
# all_modelling <- list()
# for(id in unique(dat$human_id)) {
#   all_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   # level 1
#   all_modelling[[id]]$level1_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level1")
#   all_modelling[[id]]$level1_3 <- optim(my_logit(rep(all_modelling[[id]]$level1_1$minimum,3)),fn=modelLL,data=tdat,strategy="level1")
#   # level 2
#   all_modelling[[id]]$level2_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level2")
#   all_modelling[[id]]$level2_3 <- optim(my_logit(rep(all_modelling[[id]]$level2_1$minimum,3)),fn=modelLL,data=tdat,strategy="level2")
#   # level 3
#   all_modelling[[id]]$level3_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="level3")
#   all_modelling[[id]]$level3_3 <- optim(my_logit(rep(all_modelling[[id]]$level3_1$minimum,3)),fn=modelLL,data=tdat,strategy="level3")
#   # WSLS
#   all_modelling[[id]]$WSLS_1 <- optimize(modelLL,interval=c(0,1),data=tdat,strategy="WSLS")
#   all_modelling[[id]]$WSLS_3 <- optim(my_logit(rep(all_modelling[[id]]$WSLS_1$minimum,3)),fn=modelLL,data=tdat,strategy="WSLS")
# }
# 
# save(all_modelling,file="all_modelling.RData")

```

```{r}
# create a table with BIC
# BIC_results <- data.frame()
# for(id in unique(dat$human_id)) {
#     BIC_results <- rbind(BIC_results,
#                          data.frame(
#                            "id" = id,
#                            "condition" = dat[dat$human_id==id,"condition"][1],
#                            "random_BIC" = -2*(100*log(1/3) + 50*log(1/5)), 
#                            "level1_1_theta" = all_modelling[[id]]$level1_1$minimum,
#                            "level1_1_BIC" = all_modelling[[id]]$level1_1$objective + 1*log(150),
#                            "level1_3_theta1" = my_logistic(all_modelling[[id]]$level1_3$par[1]),
#                            "level1_3_theta2" = my_logistic(all_modelling[[id]]$level1_3$par[2]),
#                            "level1_3_theta3" = my_logistic(all_modelling[[id]]$level1_3$par[3]),
#                            "level1_3_BIC" = all_modelling[[id]]$level1_3$value + 3*log(150),
#                            "level2_1_theta" = all_modelling[[id]]$level2_1$minimum,
#                            "level2_1_BIC" = all_modelling[[id]]$level2_1$objective + 1*log(150),
#                            "level2_3_theta1" = my_logistic(all_modelling[[id]]$level2_3$par[1]),
#                            "level2_3_theta2" = my_logistic(all_modelling[[id]]$level2_3$par[2]),
#                            "level2_3_theta3" = my_logistic(all_modelling[[id]]$level2_3$par[3]),
#                            "level2_3_BIC" = all_modelling[[id]]$level2_3$value + 3*log(150),
#                            "level3_1_theta" = all_modelling[[id]]$level3_1$minimum,
#                            "level3_1_BIC" = all_modelling[[id]]$level3_1$objective + 1*log(150),
#                            "level3_3_theta1" = my_logistic(all_modelling[[id]]$level3_3$par[1]),
#                            "level3_3_theta2" = my_logistic(all_modelling[[id]]$level3_3$par[2]),
#                            "level3_3_theta3" = my_logistic(all_modelling[[id]]$level3_3$par[3]),
#                            "level3_3_BIC" = all_modelling[[id]]$level3_3$value + 3*log(150),
#                            "WSLS_1_theta" = all_modelling[[id]]$WSLS_1$minimum,
#                            "WSLS_1_BIC" = all_modelling[[id]]$WSLS_1$objective + 1*log(150),
#                            "WSLS_3_theta1" = my_logistic(all_modelling[[id]]$WSLS_3$par[1]),
#                            "WSLS_3_theta2" = my_logistic(all_modelling[[id]]$WSLS_3$par[2]),
#                            "WSLS_3_theta3" = my_logistic(all_modelling[[id]]$WSLS_3$par[3]),
#                            "WSLS_3_BIC" = all_modelling[[id]]$WSLS_3$value + 3*log(150)
#                          ))
# }
# write.csv(BIC_results,file="BIC_result.csv",row.names = FALSE)
# kable(BIC_results)
```

```{r}

# create a table with best fitting models by condition
# table(dat$condition[dat$round==1 & dat$game == "rps"],c("random","level1_1","level2_1","level3_1","WSLS_1","level1_3","level2_3","level3_3","WSLS_3")[apply(BIC_results[,c("random_BIC","level1_1_BIC","level2_1_BIC","level3_1_BIC","WSLS_1_BIC","level1_3_BIC","level2_3_BIC","level3_3_BIC","WSLS_3_BIC")],1,which.min)])
# 
# ## treat 1 theta and 3 theta versions the same
# # create a table with best fitting models by condition
# table(dat$condition[dat$round==1 & dat$game == "rps"],c("random","level1","level2","level3","WSLS","level1","level2","level3","WSLS")[apply(BIC_results[,c("random_BIC","level1_1_BIC","level2_1_BIC","level3_1_BIC","WSLS_1_BIC","level1_3_BIC","level2_3_BIC","level3_3_BIC","WSLS_3_BIC")],1,which.min)])
#

```

## Fixed type: Likelihood of the different models by trial

This plot ks based on the three parameter versions of each strategy, and shows the average probability according to each model of participants' choices. It can be seen that, compared to the RPS game, where the fit grows slowly over trials, the fit in the later games increases much more rapidly.

```{r}
# long_likelihood_by_trial <- data.frame()
# for(id in unique(dat$human_id)) {
#   tdat <- subset(dat,human_id == id)
#   for(strategy in c("level1","level2","level3","WSLS")) {
#     long_likelihood_by_trial <- rbind(long_likelihood_by_trial,
#                                       data.frame(id = id,
#                                                  condition = dat[dat$human_id==id,"condition"][1],
#                                                  strategy = strategy,
#                                                  round = 1:50,
#                                                  game = rep(c("rps","fwg","numbers"),each=50),
#                                                  likelihood = modelLL(all_modelling[[id]][[paste(strategy,"_3",sep="")]]$par,
#                                                                       data=tdat,
#                                                                       strategy = strategy,
#                                                                       return_value = "likelihood_by_trial")
#                                       )
#                                 )
#   }
# }
# library(ggplot2)
# plot_dat <- long_likelihood_by_trial %>% group_by(condition,game,round,strategy) %>%
#       summarize(probability = mean(likelihood))
# # make sure the different games are ordered in the way they were played
# plot_dat$game <- factor(plot_dat$game,levels=c("rps","fwg","numbers"))
# ggplot(plot_dat,aes(x=round,y=probability,colour=strategy)) + geom_line() + facet_grid(game~condition)
```

## Fitting simple Bayesian learning model with flat prior
```{r}

best_response <- function(game,ai_action){
  if (game == "rps"){
    if (ai_action == "R") {return("P")}
    else if (ai_action == "P") {return("S")}
    else if (ai_action == "S") {return("R")}
  } else if (game == "fwg"){ 
    if (ai_action == "F") {return("W")}
    else if (ai_action == "W") {return("G")}
    else if (ai_action == "G") {return("F")}
  } else if (game == "numbers") {
    if (ai_action == "1") {return("2")}
    else if (ai_action == "2") {return("3")}
    else if (ai_action == "3") {return("4")}
    else if (ai_action == "4") {return("5")}
    else if (ai_action == "5") {return("1")}
  }
}
```

```{r}
naive_bayes <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {
  # theta = parameter, probability computer will play its best response to what it thinks human is. Truth = 90%. in [0,1].
  # Epsilone: parameter controlling noise in human choice. Probability human will deviate from its best response. In [0,1].
  # dat = data subset for one participant
  # opp_strategy_vec = vector of possible opponent strtegies, model assumes humans restrict opp strategy space to vector   c("level0","level1","level2")
  # return = -2logLik 
  # or "likelihood_by_trial" vector  (for later plotting, etc)
  # opp_mod_transfer is a boolean. True if distribution of opponent strategies is kept across games. False otherwise.

  #return_value <- match.arg(return_value)
  theta <- par[1]
  eps <- par[2]

  num_strat = length(opp_strategy_vec)

  #lik_opp is a vector that holds and update probability distribution of opp strategy given actions.
  lik_opp <- matrix(0.0,num_strat, nrow(data))
  # Initiate prior vector as uniform
  opp_prior_vec <-rep(1/num_strat, num_strat)

  lik_hum <- matrix(0.0,nrow(data))
  # br_hum is vector that stores best responses of human to actions taken by each level-k ai opponent at the round.
  br_hum <- rep(NA,num_strat)

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred <- NA
      lik_hum[t] <- 1/nopts
      if (opp_mod_transfer) {
        opp_prior_vec <- rep(1/num_strat, num_strat)
        lik_opp[,t] <- opp_prior_vec
        }

    # Rounds after first
    } else {
      # get prediction of opponent action from CSV files
      for (strategy in opp_strategy_vec) {
        k = match(strategy,opp_strategy_vec)
        pred_opp <- as.character(filter(pred_file_opp,human_previous == as.character(data[t-1,"h_action"]) & computer_previous == as.character(data[t-1,"a_action"]))[strategy][1,])
        # Given opponent action predicted, what would be human action best response for each opp strat
        br_hum[[k]] <- best_response(t_game,as.character(pred_opp))
        #cat(as.character(br_hum[[k]]), "\n")

        # Multiply prior by likelihood of observation to get posterior. Done here ot take advantage of For loop. 
        if(as.character(data[t,"a_action"]) == pred_opp) {
          lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
        } else {
          lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
        }
      }
      # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
      indices <- which(br_hum %in% as.character(data[t,"h_action"]))
      #cat(as.character(indices),  "\n")

      # THEN likelihood of current human action is just prior on opponent vec (haven't updated priors with curr opp act yet)
      # if action is not predicted by any level-k OM, assume human chooses randomly with prob eps
      if(length(indices) == 0){
        lik_hum[t] <- eps/nopts
      # if same br action for multiple possible ai opponents, add probabilities 
      } else {
        lik_hum[t] <- sum(opp_prior_vec[indices])*(1-eps) + eps/nopts
      }
      # Standardising the probabilities and updating prior
      lik_opp[,t] <- lik_opp[,t]/sum(lik_opp[,t])
      opp_prior_vec <- lik_opp[,t]
    }
    # for debugging uncomment the line below
    #cat(as.character(pred_opp),as.character(data[t,"h_action"]),lik[,t],"\n")
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)
}

```

```{r}
# 
# naive_bayes_modelling <- list()
# for(id in unique(dat$human_id)) {
#   naive_bayes_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)

#   naive_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = 0.0, upper = 1.0, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE,parallelType=1))

#   naive_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
# }
# save(naive_bayes_modelling,file="naive_bayes_modelling.RData")

# Fitting a bayesiam model with No Transfer of opp strat distribution between games 
NT_bayes_modelling <- list()
for(id in unique(dat$human_id)) {
  NT_bayes_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # NT_bayes_modelling[[id]] <- DEoptim(fn=naive_bayes, lower = 0.0, upper = 1.0, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE,parallelType=1))

  NT_bayes_modelling[[id]] <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

save(NT_bayes_modelling,file="NT_bayes_modelling.RData")

# ```

# ```{r}
# Nbayes_par_results <- data.frame()
# for(id in unique(dat$human_id)) {
# Nbayes_par_results <- rbind(Nbayes_par_results,
#                        data.frame(
#                          "id" = id,
#                          "condition" = dat[dat$human_id==id,"condition"][1],
#                          # Theta is the parameter governing AI stochasticity. Truth is 0.9
#                          "theta" = naive_bayes_modelling[[id]]$par[1],
#                          # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
#                          "eps" = naive_bayes_modelling[[id]]$par[2]))
# }
# 
# # Plotting the theta/eps scatter for all participants...
# plot(Nbayes_par_results$theta,Nbayes_par_results$eps,xlab="Theta", ylab= "Epsilone",col = ifelse(as.character(Nbayes_par_results$condition) == "Level2",'red','green'))
# 
# write.csv(Nbayes_par_results,file="Nbayes_par_results.csv",row.names = FALSE)
#kable(Nbayes_par_results)
```

## Q-learning
```{r}
Q_learn <- function(par,data,return_value=c("-2loglik","likelihood_by_trial")){
  # actions
  # beta= inverse temperature parameter in softmax choice function
  # lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]
  Q_vals_RPS = matrix(0.0,3)
  names(Q_vals_RPS) <- c("R","P","S")
  Q_vals_FWG = matrix(0.0,3)
  names(Q_vals_FWG) <- c("F","W","G")
  Q_vals_NUM = matrix(0.0,5)
  names(Q_vals_NUM) <- c("1","2","3","4","5")

  lik_hum <- matrix(0.0,nrow(data))

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
      lik_hum[t] <- 1/nopts
    } else {
      # Get reward and past human action
      reward <- as.numeric(data[t-1,"score"])
      h_act_prev <- as.character(data[t-1,"h_action"])
      Q_vals[h_act_prev] <- Q_vals[[h_act_prev]] + as.numeric(alpha)*( reward - Q_vals[[h_act_prev]])

      # cat(Q_vals,"\n")

      # Assume human chooses action probabilistically using softmax on Q values
      probs <- exp(Q_vals/beta)/sum(exp(Q_vals/beta))

      # Get actual human action and compute likelihood
      h_act <- as.character(data[t,"h_action"])
      act_index <- match(h_act, names(Q_vals))
      lik_hum[t] <- probs[[act_index]]
    }
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)

}

```

##Q-learning fit
```{r}
QL_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_modelling[[id]] <- optim(c(1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")
}

save(QL_modelling,file="QL_modelling.RData")

```

## Parametric Experience Weighted Attraction model (Camerer & Ho 1997)
```{r}
EWA_par <- function(par,data,return_value){
  # par: vector of parameters to the EWA model
  # Data :choice data by trial for each participant
  # Returns sum of loglikelihoods or likelihood per trial
  phi <- par[1]
  delta <- par[2]
  rho <- par[3]
  lambda <- par[4]

  # Define attraction vectors for each game
  A_RPS = matrix(0.0,3)
  names(A_RPS) <- c("R","P","S")
  A_FWG = matrix(0.0,3)
  names(A_FWG) <- c("F","W","G")
  A_NUM = matrix(0.0,5)
  names(A_NUM) <- c("1","2","3","4","5")

  # # Define reward matrices from the prospective of the row player (human in our case)
  reward_RPS <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_RPS) = list(c("R", "P", "S"), c("R", "P", "S"))

  reward_FWG <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_FWG) = list(c("F", "W", "G"), c("F", "W", "G"))

  reward_NUM <- t(matrix(c(0,-1,0,0,1,1,0,-1,0,0,0,1,0,-1,0,0,0,1,0,-1,-1,0,0,1,0),nrow=5, ncol=5))
  dimnames(reward_NUM) = list(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))

  # Initiate likelihood by trial vector
  lik_hum <- matrix(0.0,nrow(data))

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
      reward <- switch(as.character(t_game),rps=reward_RPS,fwg = reward_FWG, numbers = reward_NUM)
      lik_hum[t] <- 1/nopts
      # Initiate N(0) = 1 as in Camerer and Ho 1997 paper.
      N <- 1

    } else {
      # Get reward and past human action
      h_act_prev <- as.character(data[t-1,"h_action"])
      a_act_prev <- as.character(data[t-1,"a_action"])
      for (i in 1:length(Att)) {
        action <- as.character(names(Att)[i])
        # cat("this is current strat",action,"\n")
        # cat("this is previous human action",h_act_prev,"\n")

        # Attraction vector update rule
        Att[[i]] <- (phi*N*Att[[i]]  + ( delta + (1- delta)*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (rho*N + 1)
      }
      #Update the value of N
      N <- rho*N + 1

      # Assume human chooses action probabilistically using softmax on Attraction values
      probs <- exp(lambda*Att)/sum(exp(lambda*Att))

      # Get actual human action and compute likelihood
      h_act <- as.character(data[t,"h_action"])
      act_index <- match(h_act, names(Att))
      lik_hum[t] <- probs[[act_index]]
    }
  }

 if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret) ) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)
}
```

## Fitting Parametric EWA to data 
```{r}

EWA_modelling <- list()
for(id in unique(dat$human_id)) {
  EWA_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  # EWA_modelling[[id]] <- optim(c(0.5,0.5,0.5,0.5),fn=EWA_par,gr = NULL, data=tdat,"-2loglik", lower = c(0.001,0.001,0.001,0.001), upper = c(1,0.99,0.99,1), method="L-BFGS-B")
  EWA_modelling[[id]] <- DEoptim(fn=EWA_par,lower = c(0.01,0.01,0.01,0.01), upper = c(10,1,1,10), data=tdat,"-2loglik",control=list(trace = FALSE,parallelType=1))
}

save(EWA_modelling,file="EWA_modelling.RData")

```

##  Self-Tuning EWA (Camerer & Ho 2007) 
```{r}

EWA_self <- function(par,data,return_value){

  lambda <- par[1]
  # Initiate N(0) = 1 as in Camerer and Ho 1997 paper.

  # Define attraction vectors for each game 
  A_RPS = matrix(0.0,3)
  names(A_RPS) <- c("R","P","S")
  A_FWG = matrix(0.0,3)
  names(A_FWG) <- c("F","W","G")
  A_NUM = matrix(0.0,5)
  names(A_NUM) <- c("1","2","3","4","5")

  # # Define reward matrices from the prospective of the row player (human in our case)
  reward_RPS <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_RPS) = list(c("R", "P", "S"), c("R", "P", "S"))

  reward_FWG <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_FWG) = list(c("F", "W", "G"), c("F", "W", "G"))

  reward_NUM <- t(matrix(c(0,-1,0,0,1,1,0,-1,0,0,0,1,0,-1,0,0,0,1,0,-1,-1,0,0,1,0),nrow=5, ncol=5))
  dimnames(reward_NUM) = list(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))

  # Initiate likelihood by trial vector 
  lik_hum <- matrix(0.0,nrow(data))

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
      reward <- switch(as.character(t_game),rps=reward_RPS,fwg = reward_FWG, numbers = reward_NUM)
      game_data <- subset(data, game == t_game)
      lik_hum[t] <- 1/nopts
      N <- 1.0

      R_t <- rep(0.0,nopts)
      H_t <- R_t

    } else {

      indx <- data[t,"round"]
      # Get reward and past human action
      h_act_prev <- as.character(game_data[indx-1,"h_action"])
      a_act_prev <- as.character(game_data[indx-1,"a_action"])

      # Estimate phi(t)
      R_t <- as.numeric(names(Att) == a_act_prev)
      # cat("This is R(t)",R_t,"\n")

      H_t <- (H_t*(indx-2) + R_t)/(indx-1)
      # cat("this is H(t)",H_t,"\n")

      Phi_t <- 1 - 0.5*sum((H_t-R_t)^2)
      # cat("This is Phit(t)",Phi_t,"\t")

      # Estimate vector Delta(t)

      delta_t <- as.numeric(reward[,a_act_prev] >= as.numeric(game_data[indx-1,"score"]))
      #cat("this is delta(t)",delta_t,"\n")

      for (i in 1:length(Att)) {
        action <- as.character(names(Att)[i])
        # cat("this is current strat",action,"\n")
        # cat("this is previous humna action",h_act_prev,"\n")

        # Attraction vector update rule 
        Att[[i]] <- (Phi_t*N*Att[[i]]  + ( delta_t[[i]] + (1- delta_t[[i]])*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (Phi_t*N + 1)
      }
      #Update the value of N 
      N <- Phi_t*N + 1
      #cat(Att,'\n')

      # Assume human chooses action probabilistically using softmax on Attraction values
      probs <- exp(lambda*Att)/sum(exp(lambda*Att))

      # Get actual human action and compute likelihood
      h_act <- as.character(game_data[indx,"h_action"])
      act_index <- match(h_act, names(Att))
      lik_hum[t] <- probs[[act_index]]
    }
  }

 if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)
}

```

```{r}
##Plotting self_tuning EWA likelihood function for lambda from 0 to 100

lambdas <- seq(1,10,0.01)
count = 0.0

for(id in unique(dat$human_id)) {
  if (count <= 15) {
      tdat <- subset(dat,human_id == id)
      likelihoods <- sapply(lambdas, EWA_self, tdat, "-2loglik")
      plot(lambdas,likelihoods)
      count <- count + 1
  }
}

```

## Fitting Self-Tuning EWA to data 
```{r}

EWA_self_modelling <- list()
for(id in unique(dat$human_id)) {
  EWA_self_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
#   EWA_self_modelling[[id]] <- optim(10.0,fn=EWA_self,gr = NULL, data=tdat,"-2loglik", lower = 0.0, upper = 200, method="L-BFGS-B")
# }
  EWA_self_modelling[[id]] <- DEoptim(fn=EWA_self, lower = 0.0, upper = 100.0, data=tdat, "-2loglik", control=list(trace = FALSE,parallelType=1))
}

save(EWA_self_modelling,file="EWA_self_modelling.RData")

```
```{r}
cat(EWA_self_modelling[["38VxtUSv_h6RR5-tAAA2"]]$optim$bestmem)
```

## Fitting a QL with state space consisting of last round actions 
```{r}

Q_learn_states <- function(par,data,return_value){
  # beta= inverse temperature parameter in softmax choice function
  # lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # return_value=c("-2loglik","likelihood_by_trial")
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]
  gamma <- 0.9

  #Define matrix of state spaces for each game 
  G1 <- expand.grid(c("R", "P", "S"),c("R", "P", "S"))
  states_RPS <- paste0(G1$Var1,G1$Var2)

  G2 <- expand.grid(c("F", "W", "G"), c("F", "W", "G"))
  states_FWG <- paste0(G2$Var1,G2$Var2)

  G3 <- expand.grid(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
  states_NUM <- paste0(G3$Var1,G3$Var2)

  Q_vals_RPS = matrix(0.0,9,3)
  dimnames(Q_vals_RPS) = list(states_RPS, c("R", "P", "S"))

  Q_vals_FWG = matrix(0.0,9,3)
  dimnames(Q_vals_FWG) = list(states_FWG, c("F", "W", "G"))

  Q_vals_NUM = matrix(0.0,25,5)
  dimnames(Q_vals_NUM) = list(states_NUM, c("1", "2", "3", "4", "5"))

  lik_hum <- matrix(0.0,nrow(data))

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3

    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
      state_vec <- switch(as.character(t_game),rps=states_RPS,fwg = states_FWG, numbers = states_NUM)
      lik_hum[t] <- 1/nopts
      # Randomly select prev_state and actions for first round
      state_prev  <- sample(state_vec, size = 1)
      h_act_prev <- sample(colnames(Q_vals), size =1)
      a_act_prev <- sample(colnames(Q_vals), size =1)
      reward <- 0

    } else {
      # Get reward and past human action
      reward <- as.numeric(data[t-1,"score"])
      h_act_prev <- as.character(data[t-1,"h_action"])
      a_act_prev <- as.character(data[t-1,"a_action"])
    }
    # cat(state_prev, " This is the previous state", "\n")
    new_state <- paste0(h_act_prev,a_act_prev)
    # cat(new_state,"This is new state","\n")

    # Q_learning: update rule 
    Q_vals[state_prev, a_act_prev] <- Q_vals[state_prev, a_act_prev] + alpha*( reward + gamma*max(Q_vals[new_state,])- Q_vals[state_prev, a_act_prev])

    # Assume human chooses action probabilistically using softmax on Q values
    probs <- exp(Q_vals[new_state,]/beta)/sum(exp(Q_vals[new_state,]/beta))
    #cat(probs,"\n")

    # Get actual human action and compute likelihood
    h_act <- as.character(data[t,"h_action"])
    act_index <- match(h_act, colnames(Q_vals))
    lik_hum[t] <- probs[[act_index]]

    # Update state
    state_prev <- new_state 
    #}
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)

}

```

```{r}
# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# Q_learn_states(c(0.1,0.1), data, "-2loglik")

QL_states_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_states_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_states_modelling[[id]] <- optim(c(1,0.1),fn=Q_learn_states,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")
}

save(QL_states_modelling,file="QL_states_modelling.RData")

```

## Putting results together 
```{r}
All_results <- data.frame()
for(id in unique(dat$human_id)) {
All_results <- rbind(All_results,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         "random_BIC" = -2*(100*log(1/3) + 50*log(1/5)),

                         # Naive Bayes, Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "NB_theta" = naive_bayes_modelling[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "NB_eps" = naive_bayes_modelling[[id]]$par[2],
                         "NB_BIC" = naive_bayes_modelling[[id]]$value + 2*log(150),

                         # Q-Learning 
                         # beta ->  inverse temperature parameter in softmax choice function
                         "QL_Beta" = QL_modelling[[id]]$par[1],
                         # lambda -> learning rate in QL update
                         "QL_alpha" = QL_modelling[[id]]$par[2],
                         "QL_BIC" = QL_modelling[[id]]$value + 2*log(150),

                         # Parametric EWA
                         # Phi is depreciation of past attractions
                         "EWA_Phi" = EWA_modelling[[id]]$optim$bestmem[1],
                         # Delta is weight of foregone payoffs vs actual payoffs
                         "EWA_Delta" = EWA_modelling[[id]]$optim$bestmem[2],
                         # Rho is depreciation of the experience measure N(t)
                         "EWA_Rho" = EWA_modelling[[id]]$optim$bestmem[3],
                         #Lambda is a parameter of the softmax choice function (inverse Temperature)
                         "EWA_Lambda" = EWA_modelling[[id]]$optim$bestmem[4],
                         # Parametric EWA BIC
                         "EWA_BIC" = EWA_modelling[[id]]$optim$bestval + 4*log(150),
                         "EWA_2LL" = EWA_modelling[[id]]$optim$bestval,

                         # Self-Tuning EWA (only 1 parameter)
                         #Lambda is a parameter of the softmax choice function (inverse Temperature)
                         "EWA_self_Lambda" = EWA_self_modelling[[id]]$optim$bestmem[1],
                         # Parametric EWA BIC
                         "EWA_self_2LL" = EWA_self_modelling[[id]]$optim$bestval,
                         "EWA_self_BIC" = EWA_self_modelling[[id]]$optim$bestval + 1*log(150)
                         ))
}

write.csv(All_results,file="All_results.csv",row.names = FALSE)

```