### 2019-11-26 param_model_recovery.Rmd

---
title: "Simulation_EWA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### Parameter and model recovery Q-Learning with past round as state

## Build function that will simulate a QL_states player given parameters alpha and beta, and generate actions for 50 rounds
## of each game. 
```{r}
gen_data_QL_states <- function(beta, alpha, num_rounds, gamma) {

  games <- c("rps","fwg","numbers")
  condition <- sample(c("level1","level2"),1)
  data <- data.frame()

  #Define matrix of state spaces for each game 
  G1 <- expand.grid(c("R", "P", "S"),c("R", "P", "S"))
  states_RPS <- paste0(G1$Var1,G1$Var2)

  G2 <- expand.grid(c("F", "W", "G"), c("F", "W", "G"))
  states_FWG <- paste0(G2$Var1,G2$Var2)

  G3 <- expand.grid(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
  states_NUM <- paste0(G3$Var1,G3$Var2)

  # Initialize Matrix of Q-values 
  Q_vals_RPS = matrix(-0.5,9,3)
  dimnames(Q_vals_RPS) = list(states_RPS, c("R", "P", "S"))

  Q_vals_FWG = matrix(-0.5,9,3)
  dimnames(Q_vals_FWG) = list(states_FWG, c("F", "W", "G"))

  Q_vals_NUM = matrix(-0.5,25,5)
  dimnames(Q_vals_NUM) = list(states_NUM, c("1", "2", "3", "4", "5"))

  # Build tables that give reward for human player(Row) vs ai player(column) actions. 
  reward_RPS <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_RPS) = list(c("R", "P", "S"), c("R", "P", "S"))

  reward_FWG <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_FWG) = list(c("F", "W", "G"), c("F", "W", "G"))

  reward_NUM <- t(matrix(c(0,-1,0,0,1,1,0,-1,0,0,0,1,0,-1,0,0,0,1,0,-1,-1,0,0,1,0),nrow=5, ncol=5))
  dimnames(reward_NUM) = list(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))

   for (t_game in games){
      # Initiate dataframe
      game_data <- setNames(data.frame(matrix(ncol = 6, nrow = num_rounds)), c("condition","game", "round", "h_action","a_action","score"))

      # Initialize attraction vector, rewards + randomly select first actions
      Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
      state_vec <- switch(as.character(t_game),rps=states_RPS,fwg = states_FWG, numbers = states_NUM)
      game_actions <- switch(as.character(t_game),rps=c("R", "P", "S"),fwg = c("F", "W", "G"), numbers = c("1", "2", "3", "4", "5"))
      reward <- switch(as.character(t_game), rps=reward_RPS, fwg=reward_FWG, numbers=reward_NUM)
      pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)

      for(t in 1:num_rounds) {
        if(t_game == "numbers") nopts <- 5 else nopts <- 3
        game_data[t,"game"]<- t_game
        game_data[t,"round"] <- t
        game_data[t,"condition"] <- condition

        if(t == 1) {
          # Initialize previous state randomly. 
          prev_state  <- sample(state_vec, size = 1)
          # Sample huamn and ai actions in first round randomly 
          h_act <- as.character(sample(game_actions, 1 ))
          ai_act  <- as.character(sample(game_actions, 1 ))
          # Get reward from first round. 
          R <- reward[h_act, ai_act]

          #Fill in table 
          game_data[t,"h_action"] <- h_act
          game_data[t,"a_action"] <- ai_act 
          game_data[t,"score"] <- R

        } else {
          h_act<- as.character(game_data[t-1,"h_action"])
          ai_act <- as.character(game_data[t-1,"a_action"])
          R <- as.numeric(game_data[t-1,"score"])

        }

        # cat("round", t, "\n")
        # cat(prev_state, " This is the previous state", "\n")
        # cat("observed human action",h_act, "\n")
        # cat("observed ai action",ai_act, "\n")

        ###### New state is the state we get to after human plays h_act and ai plays  ai_act 
        new_state <- paste0(h_act,ai_act)
        ######
        # cat(new_state,"This is new state","\n")
        # cat("this is alpha",alpha,"\n")

        # Q_learning: update rule
        # cat("this is the Q vals",Q_vals[prev_state, h_act],"\n")
        # cat("max Q_vals",max(Q_vals[new_state,]),"\n")
        # cat("This is the reward", R ,"\n")

        Q_vals[prev_state, h_act] <- Q_vals[prev_state, h_act] + alpha*( R + gamma*max(Q_vals[new_state,]) - Q_vals[prev_state, h_act])
        probs <- exp(Q_vals[new_state,]/beta)/sum(exp(Q_vals[new_state,]/beta))

        # Select human action with softmax
        game_data[t,"h_action"] <- sample(game_actions, size=1,prob=probs)

        # Update state
        prev_state <- new_state

         # Simulate opponent action and get score 
        game_data[t,"a_action"] <- as.character(filter(pred_file_opp,human_previous == h_act & computer_previous == ai_act)[condition][1,])
        # Calculate score 
        game_data[t,"score"] <- reward[game_data[t,"h_action"], game_data[t,"a_action"]]

      }# End for loop over rounds
      data <- rbind(data,game_data)
  }
  # end of game for loop
  return(data)

}

# QL_dta <- gen_data_QL_states(5,0.5, num_rounds = 5, gamma = 0.9)

```

# Function to fit various models to simulated QL learner data and get table of results.
```{r}

fit_models <- function(tdat,num_rounds =50) {

  optim_S_EWA  <- DEoptim(fn=EWA_self, lower = 0.0, upper = 10.0, data=tdat, "-2loglik", control=list(trace = FALSE,parallelType=1))

  optim_EWA    <- DEoptim(fn=EWA_par,lower = c(0,0,0,0), upper = c(20,1,1,20), data=tdat,"-2loglik",control=list(trace = FALSE,parallelType=1))

  #optim_QL     <- optim(c(1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")

  optim_QL_states <- DEoptim(fn=Q_learn_states,lower = c(0,0), upper = c(10,1), data=tdat,"-2loglik", gamma = 0, control=list(trace = FALSE,parallelType=1))

  optim_NB     <- optim(c(0.1,0.1),fn=naive_bayes,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE, lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")

  dataframe <-  data.frame(

  # Get condition
   "condition"  = as.character(tdat$condition[1]),

  # Get best fit parameters and -2 Log Likelihoods for all models

  # Self-tuning EWA 
   "S_EWA_infrd_lambda" = optim_S_EWA$optim$bestmem[1],
   "S_EWA_2LL"          = optim_S_EWA$optim$bestval,

  #Parametric EWA
    "EWA_infrd_phi"    = optim_EWA$optim$bestmem[1],
    "EWA_infrd_delta"  = optim_EWA$optim$bestmem[2],
    "EWA_infrd_rho"    = optim_EWA$optim$bestmem[3],
    "EWA_infrd_lambda" = optim_EWA$optim$bestmem[4],
    "EWA_2LL"          = optim_EWA$optim$bestval,

  # Q-Learning with state as prev round
    "QL_states_infrd_beta"   = optim_QL_states$optim$bestmem[1],
    "QL_states_infrd_alpha"  = optim_QL_states$optim$bestmem[2],
    "QL_states_2LL"          = optim_QL_states$optim$bestval,

  # Naive Bayes (nopts-arm bandit with prior vec transfered across games)
    "NB_infrd_theta" = optim_NB$par[1],
    "NB_infrd_eps"   = optim_NB$par[2],
    "NB_2LL"         = optim_NB$value,

  # Random
   "Random_2LL"     = -2*(2*num_rounds*log(1/3) + num_rounds*log(1/5))

  )

  return(dataframe)

}

sim_results <- function(beta,alpha,num_rounds,gamma) {
  #Generate simulated data using Q_learning with previous round as state
  tdat = gen_data_QL_states(beta,alpha, num_rounds, gamma)
  # Fit various models to generated data 
  df <- fit_models(tdat)
  df$alpha <- alpha
  df$beta <- beta
  return(df)
}

```

## Run simulation to generate datasets and return results of model fits 
```{r}

betas = runif(100,0,10)
alphas = runif(100,0,1)

pars <- as.array(mapply(c,betas, alphas, SIMPLIFY = FALSE ))

list_of_dataframes <- mapply(FUN = sim_results, betas, alphas,num_rounds = 50.0, gamma = 0.0, SIMPLIFY = FALSE)

QL_states_mod_recovery <- bind_rows(list_of_dataframes, .id = "column_label")

save(QL_states_mod_recovery,file="QL_states_mod_recovery.RData")

```

## Plot infered vs real params, summarise model fitting results and return table of best fitting models
```{r}
# PLot real vs infered QL_states parameters
plot(QL_states_mod_recovery$alpha,QL_states_mod_recovery$QL_states_infrd_alpha, xlab = "True alpha", ylab = " Infered alpha")
plot(QL_states_mod_recovery$beta,QL_states_mod_recovery$QL_states_infrd_beta, xlab = "True beta", ylab = " Infered beta")

# Evaluate the correlation between real and infered alphas/lambdas and test it for significance
cor.test(QL_states_mod_recovery$alpha,QL_states_mod_recovery$QL_states_infrd_alpha, method= "spearman")
cor.test(QL_states_mod_recovery$beta,QL_states_mod_recovery$QL_states_infrd_beta, method= "spearman")

# Model recovery: create a table with best fitting models by condition
QL_states_model_recov <- table(QL_states_mod_recovery[, "condition"],c("random","N_Bayes","QL_states","EWA","S_EWA")[apply(QL_states_mod_recovery[,c("Random_2LL","NB_2LL","QL_states_2LL","EWA_2LL","S_EWA_2LL")],1,which.min)])

 write.csv(QL_states_model_recov,file="QL_states_model_recov.csv",row.names = TRUE)
 kable(QL_states_model_recov)

```

## What if we increase num rounds to 200, does that improve model recovery? 
```{r}
betas = runif(100,0,10)
alphas = runif(100,0,1)

pars <- as.array(mapply(c,betas, alphas, SIMPLIFY = FALSE ))

list_of_dataframes200 <- mapply(FUN = sim_results, betas, alphas,num_rounds = 200.0, gamma = 0.0, SIMPLIFY = FALSE)

QL_states_recov_200 <- bind_rows(list_of_dataframes200, .id = "column_label")

save(QL_states_recov_200,file="QL_states_recov_200.RData")

# PLot real vs infered QL_states parameters
plot(QL_states_recov_200$alpha,QL_states_recov_200$QL_states_infrd_alpha, xlab = "True alpha", ylab = " Infered alpha")
plot(QL_states_recov_200$beta,QL_states_recov_200$QL_states_infrd_beta, xlab = "True beta", ylab = " Infered beta")

# Evaluate the correlation between real and infered alphas/lambdas and test it for significance
cor.test(QL_states_recov_200$alpha,QL_states_recov_200$QL_states_infrd_alpha, method= "spearman")
cor.test(QL_states_recov_200$beta,QL_states_recov_200$QL_states_infrd_beta, method= "spearman")

# Model recovery: create a table with best fitting models by condition
QL_recov200_tab <- table(QL_states_recov_200[, "condition"],c("random","N_Bayes","QL_states","EWA","S_EWA")[apply(QL_states_recov_200[,c("Random_2LL","NB_2LL","QL_states_2LL","EWA_2LL","S_EWA_2LL")],1,which.min)])

 write.csv(QL_recov200_tab,file="QL_recov200_tab",row.names = TRUE)
 kable(QL_recov200_tab)
```

### Model and parameter recovery for Self Tuning EWA model

## Build function that will simulate a self tuning EWA player given parameter Lambda, and generate actions for 50 rounds
# of each game. 
```{r}

# NOTE : haven't introduced noise in either human or computer actions.

gen_data_SEWA <- function(par,num_rounds){

  lambda <- par[1]
  games <- c("rps","fwg","numbers")
  condition <- sample(c("level1","level2"),1)
  data <- data.frame()

  # Define attraction vectors for each game 
  A_RPS = matrix(0.0,3)
  names(A_RPS) <- c("R","P","S")
  A_FWG = matrix(0.0,3)
  names(A_FWG) <- c("F","W","G")
  A_NUM = matrix(0.0,5)
  names(A_NUM) <- c("1","2","3","4","5")

  # # Define reward matrices from the prospective of the row player (human in our case)
  reward_RPS <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_RPS) = list(c("R", "P", "S"), c("R", "P", "S"))

  reward_FWG <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
  dimnames(reward_FWG) = list(c("F", "W", "G"), c("F", "W", "G"))

  reward_NUM <- t(matrix(c(0,-1,0,0,1,1,0,-1,0,0,0,1,0,-1,0,0,0,1,0,-1,-1,0,0,1,0),nrow=5, ncol=5))
  dimnames(reward_NUM) = list(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))

  for (t_game in games){
      # Initiate dataframe
      game_data <- setNames(data.frame(matrix(ncol = 6, nrow = num_rounds)), c("condition","game", "round", "h_action","a_action","score"))
      # Initialize attraction vector, rewards + randomly select first actions
      Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
      game_actions <- names(Att)
      reward <- switch(as.character(t_game), rps=reward_RPS, fwg=reward_FWG, numbers=reward_NUM)
      pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)

    for(t in 1:num_rounds) {
      if(t_game == "numbers") nopts <- 5 else nopts <- 3
      game_data[t,"game"]<- t_game
      game_data[t,"round"] <- t
      game_data[t,"condition"] <- condition

      if(t == 1) {

        game_data[t,"h_action"] <- as.character(sample(game_actions, 1 ))
        game_data[t,"a_action"] <- as.character(sample(game_actions, 1 ))
        game_data[t,"score"] <- reward[game_data[t,"h_action"], game_data[t,"a_action"]]
        lik_hum <- c(1/nopts)
        N <- 1.0
        R_t <- rep(0.0,nopts)
        H_t <- R_t

      } else {

        # Get reward and past human action
        h_act_prev <- as.character(game_data[t-1,"h_action"])
        a_act_prev <- as.character(game_data[t-1,"a_action"])

        # Estimate phi(t)
        R_t <- as.numeric(game_actions == a_act_prev)
        # cat("This is R(t)",R_t,"\n")

        H_t <- (H_t*(t-2) + R_t)/(t-1)
        # cat("this is H(t)",H_t,"\n")

        Phi_t <- 1 - 0.5*sum((H_t-R_t)^2)
        # cat("This is Phit(t)",Phi_t,"\t")

        # Estimate vector Delta(t)
        delta_t <- as.numeric(reward[,a_act_prev] >= as.numeric(game_data[t-1,"score"]))
        #cat("this is delta(t)",delta_t,"\n")

        for (i in 1:length(Att)) {
          action <- as.character(game_actions[i])
          # Attraction vector update rule 
          Att[[i]] <- (Phi_t*N*Att[[i]]  + ( delta_t[[i]] + (1- delta_t[[i]])*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (Phi_t*N + 1)
        }
        #Update the value of N 
        N <- Phi_t*N + 1
        #cat(Att,'\n')

        # Assume human chooses action probabilistically using softmax on Attraction values
        probs <- exp(lambda*Att)/sum(exp(lambda*Att))

        # Get actual human action and compute likelihood
        game_data[t,"h_action"] <- sample(game_actions, size=1,prob=probs)
        act_index <- match(game_data[t,"h_action"], game_actions)
        lik_hum[t] <- probs[[act_index]]

        # Simulate opponent action and get score 
        game_data[t,"a_action"] <- as.character(filter(pred_file_opp,human_previous == as.character(game_data[t-1,"h_action"]) & computer_previous == as.character(game_data[t-1,"a_action"]))[condition][1,])
        game_data[t,"score"] <- reward[game_data[t,"h_action"], game_data[t,"a_action"]]

      }

    }# End for loop over rounds
    data <- rbind(data,game_data)

  }# end of game for loop
  return(data)

}

```

## Generate 100 datasets using S_EWA agent, one for each lambda and fit the 4 models to it : NB, QL, EWA and S_EWA 
```{r}
## Generate N lambdas in chosen interval, get the corresponding datasets of actions and scores, use Self_EWA to infer Lambda from these datasets. 
EWA_self_simulation <- data.frame()
lambdas = runif(100, 0, 10)
num_rounds <- 50 
count <- 1

for(lambda in lambdas) {

  EWA_self_simulation[count,"real_lambda"] <- lambda
  tdat <- gen_data_SEWA(par = lambda, num_rounds)

  # Fit various models to generated data (from S_EWA type player)
  optim_S_EWA  <- DEoptim(fn = EWA_self, lower = 0.0, upper = 10.0, data= tdat, "-2loglik", control = list(trace = FALSE,parallelType =1))

  optim_EWA    <- DEoptim(fn = EWA_par,lower = c(0,0,0,0), upper = c(20,1,1,20), data=tdat,"-2loglik",control= list(trace = FALSE,parallelType=1))

  # optim_QL     <- optim(c(1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0,0), upper = c(10,0.99), method="L-BFGS-B")

  optim_QL_states <- DEoptim(fn = Q_learn_states,lower = c(0,0), upper = c(10,1), data=tdat,"-2loglik", gamma = 0, control=list(trace = FALSE,parallelType=1))

  optim_NB     <- optim(c(0.1,0.1),fn = naive_bayes, gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", lower = c(0.01,0.01), upper = c(0.99,0.99), method ="L-BFGS-B")

  # Get condition
  EWA_self_simulation[count, "condition"]  <- as.character(tdat$condition[1])

  # Get best fit parameters and -2 Log Likelihoods for all models

  # Self-tuning EWA 
  EWA_self_simulation[count, "S_EWA_infrd_lambda"] <- optim_S_EWA$optim$bestmem[1]
  EWA_self_simulation[count, "S_EWA_2LL"]          <- optim_S_EWA$optim$bestval

  #Parametric EWA
  EWA_self_simulation[count, "EWA_infrd_phi"]    <- optim_EWA$optim$bestmem[1]
  EWA_self_simulation[count, "EWA_infrd_delta"]  <- optim_EWA$optim$bestmem[2]
  EWA_self_simulation[count, "EWA_infrd_rho"]    <- optim_EWA$optim$bestmem[3]
  EWA_self_simulation[count, "EWA_infrd_lambda"] <- optim_EWA$optim$bestmem[4]
  EWA_self_simulation[count, "EWA_2LL"]          <- optim_EWA$optim$bestval

  # Q-Learning with previous round as state
  EWA_self_simulation[count, "QL_infrd_beta"]   <- optim_QL_states$optim$bestmem[1]
  EWA_self_simulation[count, "QL_infrd_alpha"]  <- optim_QL_states$optim$bestmem[2]
  EWA_self_simulation[count, "QL_2LL"]          <- optim_QL_states$optim$bestval

  # Naive Bayes (nopts-arm bandit with prior vec transfered across games)
  EWA_self_simulation[count, "NB_infrd_theta"] <- optim_NB$par[1]
  EWA_self_simulation[count, "NB_infrd_eps"]   <- optim_NB$par[2]
  EWA_self_simulation[count, "NB_2LL"]         <- optim_NB$value

  # Random
  EWA_self_simulation[count, "Random_2LL"]     <- -2*(2*num_rounds*log(1/3) + num_rounds*log(1/5))

  count <- count + 1
}

save(EWA_self_simulation,file="SEWA_simulation.RData")

```

## Plotting parameter recovery scatter plots and model recovery results
```{r}

plot(EWA_self_simulation$real_lambda,EWA_self_simulation$S_EWA_infrd_lambda)
# We can evaluate the correlation between real and infered lambdas and test it for significance
cor.test(EWA_self_simulation$real_lambda, EWA_self_simulation$S_EWA_infrd_lambda, method=c("spearman"))

# Model recovery: create a table with best fitting models by condition
SEWA_recov_results <- table(EWA_self_simulation[, "condition"],c("random","N_Bayes","QL","EWA","S_EWA")[apply(EWA_self_simulation[,c("Random_2LL","NB_2LL","QL_2LL","EWA_2LL","S_EWA_2LL")],1,which.min)])

 write.csv(SEWA_recov_results,file="SEWA_recov_results.csv",row.names = TRUE)
 kable(SEWA_recov_results)
```

```{r}

```