### 2019-12-04 Experiment_2/exp2_score_plots.Rmd

---
title: "Cumulative Score Plots Experiment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
```

```{r}

dat_exp2 = read.csv("data_exp2.csv")

#Adjust shootout score tor eflect easier game
dat_exp2$adj_score <- ifelse(as.character(dat_exp2$game) == "shootout", dat_exp2$score - 0.333, dat_exp2$score)

# Load data comparing models BIC for each participant 
model_comp = read.csv("exp2_model_comp.csv")
```

```{r}
datalist2 = list()
i = 0

#Build empty datafrsme with same names as original data
exp2_cum_score <- setNames(data.frame(matrix(ncol = ncol(dat_exp2), nrow = 0)), colnames(dat_exp2))
for(id in unique(dat_exp2$human_id)) {
  i <- i+1
  tdat <- subset(dat_exp2,human_id == id)
  tdat$part_num <- i
  # NB : USE ADJUSTED SCORE IN SHOOTOUT
  tdat <- within(tdat, acc_sum <- cumsum(tdat$adj_score))
  datalist2[[i]] <- tdat
}

# Merge all datasets into one 
exp2_cum_score <- dplyr::bind_rows(datalist2)
# or exp2_cum_score <- data.table::rbindlist(datalist)

# Add column for time t
exp2_cum_score <- exp2_cum_score %>% group_by(human_id) %>% mutate(t = row_number())

# Participant number as a factor 
# tdat$part_num <- as.factor(tdat$part_num)

# Add best fitting model per participant
exp2_cum_score <- merge(exp2_cum_score, model_comp[, c("human_id", "Best_model")], by="human_id")
```

```{r}
ggplot(data = exp2_cum_score, aes(x = t, y=acc_sum, group = part_num)) + 
  geom_line(aes(color= Best_model))
```

```{r}
temp <- exp2_cum_score[,c("t","acc_sum","Best_model","condition","part_num")]
dat_by_model <- temp %>% group_by(Best_model,t) %>% 
  summarize(model_acc_sum = mean(acc_sum))

ggplot(data = dat_by_model, aes(x = t, y=model_acc_sum, group = Best_model)) + 
   geom_line(aes(color= Best_model)) +
  scale_x_continuous(minor_breaks = seq(0 , 180, 10), breaks = seq(0, 180, 60))
```