### 2019-12-04 score_plots.Rmd

---
title: "Score_plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

exp1_dat = read.csv("exp1_data.csv")
exp1_model_comp = read.csv("exp1_model_comp.csv")

datalist = list()
i = 0
new_dat <- setNames(data.frame(matrix(ncol = ncol(exp1_dat), nrow = 0)), colnames(exp1_dat))
for(id in unique(exp1_dat$human_id)) {
  i <- i+1
  tdat <- subset(exp1_dat,human_id == id)
  tdat$part_num <- i
  tdat <- within(tdat, acc_sum <- cumsum(tdat$score))
  datalist[[i]] <- tdat
}

# Merge all datasets into one 
new_dat <- dplyr::bind_rows(datalist)
# or new_dat <- data.table::rbindlist(datalist)

# Add column for time t
new_dat <- new_dat %>% group_by(exp1_dat$human_id) %>% mutate(t = row_number())

# Participant number as a factor 
# tdat$part_num <- as.factor(tdat$part_num)

# Add best fitting model per participant
new_dat <- merge(new_dat, exp1_model_comp[, c("human_id", "Best_model")], by="human_id")
```

```{r}
ggplot(data = new_dat, aes(x = t, y=acc_sum, group = part_num)) + 
  geom_line(aes(color= Best_model))
```

```{r}
temp <- new_dat[,c("t","acc_sum","Best_model","condition","part_num")]
dat_by_model <- temp %>% group_by(Best_model,t) %>% 
  summarize(model_acc_sum = mean(acc_sum))

ggplot(data = dat_by_model, aes(x = t, y=model_acc_sum, group = Best_model)) + 
   geom_line(aes(color= Best_model))
```