### 2019-12-05 Report/draft_report.Rnw
report init

\documentclass{article}

\begin{document}
\SweaveOpts{concordance=TRUE}

\section*{Introduction}

\section*{Methods}

\section*{Results}

\section*{Discussion}

\section*{Conclusion}

\end{document}